---
title: "Homework4"
output: rmarkdown::html_vignette
author: Yiziying Chen
vignette: >
  %\VignetteIndexEntry{Homework4}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

This homework is due by the end of the day on November 5th 2018. Solutions should appear as a vignette in your package called "homework-4". 

1. CASL Number 2 in Exercises 5.8.
The Hessian matrix is written as: $H(l) = X^t \cdot diag(p(1-p)) \cdot X$. When p values are extremely close to 1 or 0, the diagnal is 0, which is not well-conditioned for logistic variation. 
Now, generate a diagnal matrix, with p values equal to 1. The matrix X is given below:
```{r}
p <- 0
prob <- p*(1-p)
X <- diag(x = prob, ncol = 10, nrow = 10, name = TRUE)
```

2. CASL Number 4 in Exercises 5.8. 
```{r}
# Inputs: X a numeric design matrix; y numeric vector;
#         family, an instance of a family object;
#         maxit, integer maximum iterations; tol, stopping
#         criterion tolerance; lambda, a fixed tunning parameter
# Output: estimate of the GLM regression vector including a ridge penalty on it
irwls_glm <- function(X, y, family, maxit=25, tol=1e-10, lambda){
  beta <- rep(0,ncol(X))
  for(j in 1:maxit){
  b_old <- beta
  eta <- X %*% beta
  mu <- family$linkinv(eta)
  mu_p <- family$mu.eta(eta)
  z <- eta + (y - mu) / mu_p
  W <- as.numeric(mu_p^2 / family$variance(mu))
  XtX <- crossprod(X, diag(W) %*% X) + 2*lambda*diag(rep(1,ncol(X)))
  Xtz <- crossprod(X, W * z)
  beta <- solve(XtX, Xtz)
  if(sqrt(crossprod(beta - b_old)) < tol) break
}
beta
}

n <- 1000; p <- 3
beta <- c(0.2, 2, 1)
X <- cbind(1, matrix(rnorm(n * (p- 1)), ncol = p - 1))
mu <- 1 - pcauchy(X %*% beta)
y <- as.numeric(runif(n) > mu)
irwls_glm(X, y, family = binomial(link = "cauchit"), lambda = 0.1)
```

3. Consider the sparse matrix implementation from class and the sparse add
`sparse_mulyiply` function:
```{r}
sparse_multiply <- function(a, b) {
  colnames(b) <- c("i2", "j2", "x2")
  c <- merge(a, b, by.x = "j", by.y = "i2",
  all = FALSE, suffixes = c("1", "2"))
  c$x <- c$x * c$x2
  c$key <- paste(c$i, c$j, sep = "-")
  x <- tapply(c$x, c$key, sum)
  key <- strsplit(names(x), "-")
  d <- data.frame(i = sapply(key, getElement, 1),
  j = sapply(key, getElement, 2),
  x = as.numeric(x))
  return(d)
}
```

```{r}
#write a sparse.matrix function that create a new class of "sparse.matrix"
sparse.matrix <- function(x, i, j, dims = NULL){
  result <- data.frame(i, j, x)
  if (is.null(dims)){
    dims<- c(max(i), max(j))
  }
  
  
  index <- data.frame(i = rep(1:dims[1]), j = rep(1:dims[2], each = dims[1]))
  result <- merge(index, result, by = c("i", "j"), all.x = TRUE, sort = FALSE)
  result1 <- result[result$x !=0 & !is.na(result$x),]
  
  attr(result1$i, "label") <- dims[1]
  attr(result1$j, "label") <- dims[2]
  class(result1) <- c("sparse.matrix", class(result1))
  result1
}

sparse.matrix(i = c(1, 2), j = c(1, 1), x = c(3, 1), dims = c(3, 2))
```

```{r}
#write %*% method inside of "sparse.matrix"
`%*%.default` <- .Primitive("%*%")

`%*%` <- function(x,...){
  UseMethod("%*%", x)
}

`%*%` <- function(x,y){
  UseMethod("%*%", x)
}


`%*%.sparse.matrix` <- function(a,b){
  if(!inherits(b, "sparse.matrix")){
    stop("b argument is not of a sparse matrix type.")
  }
  if(attr(a$j, "label") != attr(b$i, "label")){
    stop("the dimensions of 2 input matrcies don't match")
  }
  if(attr(a$i, "label") ==0 || attr(a$j, "label") ==0 || 
     attr(b$i, "label") ==0 || attr(b$j, "label") ==0 ){
    row_a <- max(a$i)
    col_a <- max(a$j)
    row_b <- max(b$i)
    col_b <- max(b$j)
  }
  else{
    row_a <- attr(a$i, "label")
    col_a <- attr(a$j, "label")
    row_b <- attr(b$i, "label")
    col_b <- attr(b$j, "label")
  }
  index <- data.frame(i = rep(1:max(row_a, row_b)), j = rep(1:max(col_a, col_b), each = max(row_a, row_b)))
  c <- merge(index, a, by = c("i", "j"), all.x = TRUE, sort = FALSE)
  c1 <- merge(c, b, by = c("i", "j"), all = TRUE, suffixes = c("1", "2"))
  c1$x1[is.na(c1$x1)]<- 0
  c1$x2[is.na(c1$x2)]<- 0
  
  for (m in 1:row_a){
    for (n in 1:col_b){
      c1$x[c1$i == m & c1$j == n] <- sum(c1$x1[c1$i == m] * c1$x2[c1$j == n])
    }
  }
  
  c2 <- c1[c1$x !=0 & !is.na(c1$x),]
  c3 <- c2[,c("i", "j", "x")]

  final<- sparse.matrix(i = c3$i, j = c3$j, x = c3$x, dims = c(attr(a$i, "label"), attr(b$j, "label")))
  final
}


sm0 <- sparse.matrix(i = c(1, 2), j = c(1, 1), x = c(1, 1))
sm1 <- sparse.matrix(i = c(1, 2), j = c(1, 1), x = c(3, 1), dims = c(3, 2))
sm2 <- sparse.matrix(i = c(1, 2, 2), j = c(1, 1, 2), x = c(4.4, 1.2, 3),
                     dims = c(2, 3))
sm3 <- sparse.matrix(i = rep(1, 3), j = 1:3, x = 1:3, dims = c(2, 3))


```




```{r}
#write a `+` method for sparse.matrix
`+.sparse.matrix` <- function(a,b){
  if(!inherits(b, "sparse.matrix")){
    stop("b argument is not of a sparse matrix type.")
  }
  if(attr(a$j, "label") != attr(b$j, "label") || attr(a$i, "label") != attr(b$i, "label")){
    stop("2 input matrcies can't be added up together")
  }
  if(attr(a$i, "label") ==0 || attr(a$j, "label") ==0 || 
     attr(b$i, "label") ==0 || attr(b$j, "label") ==0 ){
    row_a <- max(a$i)
    col_a <- max(a$j)
    row_b <- max(b$i)
    col_b <- max(b$j)
  }
  else{
    row_a <- attr(a$i, "label")
    col_a <- attr(a$j, "label")
    row_b <- attr(b$i, "label")
    col_b <- attr(b$j, "label")
  }
  
  a.1 <- data.frame(i = a$i, j = a$j, x = a$x)
  b.1 <- data.frame(i = b$i, j = b$j, x = b$x)
  index.1 <- data.frame(i = rep(1:max(row_a, row_b)), j = rep(1:max(col_a, col_b), each = max(row_a, row_b)))
  ad <- merge(index.1, a.1, by = c("i", "j"), all.x = TRUE, sort = FALSE)
  ad1 <- merge(ad, b.1, by = c("i", "j"), all = TRUE, suffixes = c("1", "2"))
  ad1$x1[is.na(ad1$x1)]<- 0
  ad1$x2[is.na(ad1$x2)]<- 0
  ad1$x <- ad1$x1 + ad1$x2
  ad2 <- ad1[ad1$x !=0 & !is.na(ad1$x),]
  ad3 <- ad2[, c("i", "j", "x")]

  ad.fin<- sparse.matrix(i = ad3$i, j = ad3$j, x = ad3$x, dims = c(max(ad3$i), max(ad3$j)))
  ad.fin
  
}

sm0 + sm0
sparse.matrix(i = c(1, 2), j = c(1, 1), x = c(2, 2), 
                             dims = c(2, 1))


sm2 + sm3
sparse.matrix(i = c(rep(1:2, 2), 1), j = c(1, 1, 2, 2, 3),
                             x = c(5.4, 1.2, 2, 3, 3))
```



```{r}
#write a `t()` method to sparse.matrix
`t.sparse.matrix` <- function(a){
   if(!inherits(a, "sparse.matrix")){
    stop("a argument is not of a sparse matrix type.")
  }
 
  if(attr(a$i, "label") ==0 || attr(a$j, "label") == 0){
    row_a <- max(a$i)
    col_a <- max(a$j)}
  else{
    row_a <- attr(a$i, "label")
    col_a <- attr(a$j, "label")}
  
  index.1 <- data.frame(i = rep(1:row_a), j = rep(1:col_a, each =row_a))
  at<- data.frame(i = a$i, j = a$j, x = a$x)
  at$new.i <- at$j
  at$new.j <- at$i
  at1 <- at[, c("new.i", "new.j", "x")]
  colnames(at1) <- c("i", "j", "x")
  at2 <- merge(index.1, at1, by = c("i", "j"), all.x = TRUE, sort = FALSE)
  at3 <- at2[at2$x !=0 & !is.na(at2$x),]
 
  attr(at3$i, "label") <- attr(a$j, "label")
  attr(at3$j, "label") <- attr(a$i, "label")
  
  class(at3) <- c("sparse.matrix", class(at3))
  at3

}

```






