---
title: "BIS557 Homework-5 Neural Networks"
output: rmarkdown::html_vignette
author: Yiziying Chen
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

1.
the code used below was taken from:
https://keras.rstudio.com/articles/examples/mnist_cnn.html

-----------------------------------------------------------------------------------------
```{r eval=FALSE}
install.packages("keras")
library(keras)
library(bis557)
# Data Preparation -----------------------------------------------------

batch_size <- 128
num_classes <- 10
epochs <- 2

# Input image dimensions
img_rows <- 28
img_cols <- 28

# The data, shuffled and split between train and test sets
mnist <- dataset_mnist()
x_train <- mnist$train$x
y_train <- mnist$train$y
x_test <- mnist$test$x
y_test <- mnist$test$y

# Redefine  dimension of train/test inputs
x_train <- array_reshape(x_train, c(nrow(x_train), img_rows, img_cols, 1))
x_test <- array_reshape(x_test, c(nrow(x_test), img_rows, img_cols, 1))
input_shape <- c(img_rows, img_cols, 1)

# Transform RGB values into [0,1] range
x_train <- x_train / 255
x_test <- x_test / 255

cat('x_train_shape:', dim(x_train), '\n')
cat(nrow(x_train), 'train samples\n')
cat(nrow(x_test), 'test samples\n')

# Convert class vectors to binary class matrices
y_train <- to_categorical(y_train, num_classes)
y_test <- to_categorical(y_test, num_classes)

# Define Model -----------------------------------------------------------

# Define model
model <- keras_model_sequential() %>%
  layer_conv_2d(filters = 32, kernel_size = c(3,3), activation = 'relu',
                input_shape = input_shape) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3,3), activation = 'relu') %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_dropout(rate = 0.25) %>% 
  layer_flatten() %>% 
  layer_dense(units = 128, activation = 'relu') %>% 
  layer_dropout(rate = 0.5) %>% 
  layer_dense(units = num_classes, activation = 'softmax')

# Compile model
model %>% compile(
  loss = loss_categorical_crossentropy,
  optimizer = optimizer_adadelta(),
  metrics = c('accuracy')
)

# Train model
model %>% fit(
  x_train, y_train,
  batch_size = batch_size,
  epochs = epochs,
  validation_split = 0.2
)

model %>% fit(
  x_test, y_test,
  batch_size = batch_size,
  epochs = epochs,
  validtaion_split = 0.2
)

scores <- model %>% evaluate(
  x_test, y_test, verbose = 0
)

l <- model %>% evaluate(
  x_train, y_train, verbose = 0
)
# Output metrics
cat('Test loss:', scores[[1]], '\n')
cat('Test accuracy:', scores[[2]], '\n')

cat('Train loss:', l[[1]], '\n')
cat('Train accuracy:', l[[2]], '\n')

----------------------------------------------------------------------------------------

sd(x_train)
mean(x_train)
kurtosis(x_train)

sd(x_test)
mean(x_test)
kurtosis(x_test)
```

The train loss and test loss values are given from the images as predictive features for accuracy.
The test loss has a value of 0.02102579 and accuracy of 0.9927; Train loss has a value of 0.05050852 and accuracy of 0.9845833.


2.
Note :because emnist_train and emnits_test have big sizes and exceed CRAN package size limit, I won't include these 2 data packages into bis557.
```{r eval=FALSE}

data("emnist_test") 
data("emnist_train")
y_test = as.matrix(emnist_test$`1`)
x_test = as.matrix(emnist_test[,-1])
y_train = as.matrix(emnist_train$`23`)
x_train = as.matrix(emnist_train[,-1])
x_train <- array_reshape(x_train, c(nrow(x_train), 28, 28, 1))
x_test <- array_reshape(x_test, c(nrow(x_test), 28, 28, 1))


model <- keras_model_sequential()
model %>% layer_conv_2d(filters = 32, kernel_size = c(4,4),
input_shape = c(28, 28, 1), padding = "same") %>%
layer_activation(activation = "relu") %>% layer_conv_2d(filters = 32, kernel_size = c(4,4),
padding = "same") %>%layer_activation(activation = "relu") %>%
layer_max_pooling_2d(pool_size = c(4, 4)) %>% layer_dropout(rate = 0.5) %>%
layer_conv_2d(filters = 32, kernel_size = c(4,4), padding = "same") %>%
layer_activation(activation = "relu") %>%
layer_conv_2d(filters = 32, kernel_size = c(4,4),
padding = "same") %>%
layer_activation(activation = "relu") %>%
layer_max_pooling_2d(pool_size = c(4, 4)) %>%
layer_dropout(rate = 0.5) %>%
layer_flatten() %>%
layer_dense(units = 128) %>%
layer_activation(activation = "relu") %>%
layer_dense(units = 128) %>%
layer_activation(activation = "relu") %>%
layer_dropout(rate = 0.5) %>%
layer_dense(units = 26) %>%
layer_activation(activation = "softmax")

model %>% compile(loss = 'categorical_crossentropy',
optimizer = optimizer_rmsprop(), metrics = c('accuracy'))

p1 <- predict_classes(model, x_train)
count1 = 0
for (i in 1:length(y_train)){
  if (y_train[i] == p1[i]){
    count1 = count1 + 1
  }
}
#kernal size changes
a1 = count1 / length(y_train)
a1.2 = 0.03837881
a1.3 = 0.02745526
a1.3.5 = 0.04377302
a1.4 = 0.0306535
#pool size changes
a2.3 = 0.01792813
a2.3.5 = 0.03743285
a2.4 = 0.02708364
a4.4 = 0.0382115
```

When changing kernal size from c(2,2) to c(4,4), the accuracy does not necessarily increase all the way, it decreases at kernal size = c(3,3) and increases at kernal size = c(3.5, 3.5), and further decreases at kernal size = c(4,4). Also, changing pool size does not necessarily increase accuracy. Compared with at pool size = c(2,2), accuracy decreases at pool size c(3,3) and increases at pool size = c(3.5, 3.5) and further decreases at pool size = c(4,4). So in order to improve classfication rate, we need to perform multiple tries to find the optimal accuracy when changing kernal size or pool size.

3.
```{r}
# Inputs: y, response vector; a, predicted
# response
# Output: gradient of the absolute deviation loss function
abs_p <- function(y, a) {
  if(a<y){
    der = -1
  }
  if(a>=y){
    der = 1
  }
der
}

# Inputs: v, a vector, matrix or array
# Output: input with negative terms pushed to zero
ReLU <- function(v) {
v[v < 0] <- 0
v
}

# Inputs: v, a vector, matrix or array
# Output: converts positive terms to one and
# non-positive terms to zero.
ReLU_p <- function(v) {
  p <- v * 0
p[v > 0] <- 1
p
}


# Inputs: sizes, a vector giving the size of
# layer in the neural network, including
# the input and output layers
# Output: a list containing initialized weights,
# biases, and momentums
make_weights <- function(sizes) {
L <- length(sizes) - 1
weights <- vector("list", L)
for (j in seq_len(L)) {
w <- matrix(rnorm(sizes[j] * sizes[j + 1],
sd = 1/sqrt(sizes[j])),
ncol = sizes[j],
nrow = sizes[j + 1])
v <- matrix(0,
ncol = sizes[j],
nrow = sizes[j + 1])
weights[[j]] <- list(w = w,
v = v,
b = rnorm(sizes[j + 1]))
}
weights
}

f_prop <- function(x, weights, sigma) {
L <- length(weights)
z <- vector("list", L)
a <- vector("list", L)
for (j in seq_len(L)) {
a_j1 <- if(j == 1) x else a[[j - 1L]]
z[[j]] <- weights[[j]]$w %*% a_j1 + weights[[j]]$b
a[[j]] <- if (j != L) sigma(z[[j]]) else z[[j]]
}
return(list(z = z, a = a))
}

b_prop <- function(x, y, weights, f_obj, sigma_p, f_p) {
z <- f_obj$z; a <- f_obj$a
L <- length(weights)
grad_z <- vector("list", L)
grad_w <- vector("list", L)
for (j in rev(seq_len(L))) {
if (j == L) {
grad_z[[j]] <- f_p(y, a[[j]])
} else {
grad_z[[j]] <- (t(weights[[j + 1]]$w) %*%
grad_z[[j + 1]]) * sigma_p(z[[j]])
}
a_j1 <- if(j == 1) x else a[[j - 1L]]
grad_w[[j]] <- grad_z[[j]] %*% t(a_j1)
}
return(list(grad_z = grad_z, grad_w = grad_w))
}


# Inputs: X, data matrix; y, response; sizes, vector
# given the number of neurons in each layer of
# the neural network; epochs, integer number of
# epochs to complete; eta, learning rate; mu,
# momentum term; l2, penalty term for l2-norm;
# weights, optional list of starting weights
# Output: the trained weights for the neural network
nn_sgd <- function(X, y, sizes, epochs, eta, mu = 0,
l2 = 0, weights = NULL) {
if (is.null(weights))
weights <- make_weights(sizes)
for (epoch in seq_len(epochs)) {
for (i in seq_len(nrow(X))) {
f_obj <- f_prop(X[i,], weights, ReLU)
b_obj <- b_prop(X[i,], y[i,], weights, f_obj,
ReLU_p, abs_p)
for (j in seq_along(b_obj)) {
weights[[j]]$b <- weights[[j]]$b -
eta * b_obj$grad_z[[j]]
weights[[j]]$v <- mu * weights[[j]]$v -
eta * b_obj$grad_w[[j]]
weights[[j]]$w <- (1 - eta * l2) *
weights[[j]]$w +
weights[[j]]$v
}
}
}
return(weights)
}

l2_norm <- rep(NA_real_, 3)
l2_vals <- c(0, 0.01, 0.04, 0.05)
X <- matrix(runif(1000, min = -1, max = 1), ncol = 1)
y <- X[,1,drop = FALSE]^2 + rnorm(1000, sd = 0.1)
weights_start <- make_weights(sizes = c(1, 10, 1))
for (i in seq_along(l2_vals)) {
weights <- nn_sgd(X, y, weights = weights_start,
epochs = 10, eta = 0.1,
l2 = l2_vals[i])
l2_norm[i] <- sum((weights[[1]]$w)^2)
}
l2_norm
```

With absolute deviation loss function derivative, we have see the size of weights decrease as the penalty term increases (11.991938,3.70796,1.45447,1.309176). Compared with using mse (6.942948e+00, 1.674410e+00, 3.672386e-02, 7.117368e-07), the decrease in weights is smaller using absolute deviation loss function.
