---
title: "Homework3"
output: rmarkdown::html_vignette
author: Yiziying Chen
vignette: >
  %\VignetteIndexEntry{Homework3}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

##Q1
Write a function `kern_density` that accepts a training vector x, bandwidth h, and test set x_new, returning the kernel density estimate from the Epanechnikov kernel. Visually test how this performs for some hand constructed datasets and bandwidths.


kernal : 1/h * K(x/h)
Epanechnikov kernel is given by 
$K(x) = \frac{3}{4} (1-x^2) . 1_{|x| \leq 1}$
```{r}
#calculate Epanechnikov kernel
kernel_epan <-function(x, h=1){
x <- x / h
ran <- as.numeric(abs(x) <= 1)
val <- (3/4) * ( 1 - x^2 ) * ran * (1/h)
return(val)
}

# calculate kernel density
# input x is a training vector; x_new is a test set, h is bandwith
# output a kernel density estimation from Epanechnikov kernel 
kern_density <- function(x, x_new, h=1){
  sapply(x_new, function(v){
  w <- kernel_epan((x - v), h=h)
  dens <- sum(w)/length(x)
  dens
})
}

#checking the visual implementation results
x <- rnorm(100)
x_new <- sort(rnorm(100))
dens1 <- kern_density(x, x_new, h=1)
plot(y = dens1, x = x_new, main = "Kernal density estimation with bandwith = 1", ylim = c(0,0.5))
curve(dnorm(x,0,1), col="darkblue", lwd=2, add=TRUE)
```

##Q2
###Show that if f and g are both convex functions, then their sum must also be convex.

For convect functions $f$ and $g$, assuming $0\leq \lambda \leq 1$, the definition of convextivity shows:

$f(\lambda x + (1-\lambda)x) \leq \lambda f(x) + (1-\lambda) f(x)$. 

So, $h = f+g$ where f and g are convex satisfies:

$h(\lambda x + (1-\lambda)x) = f(\lambda x + (1-\lambda)x) + g(\lambda x + (1-\lambda)x) \leq \lambda (f(x) + g(x)) + (1-\lambda)(f(x) + g(x))$

##Q3
###Illustrate that the absolute value function is convex. Using the result from the previous exercise, show that the `1-norm is also convex.

By definition, f is convext on real interval $I$ if and only if: $\forall x, y \in I: \forall \alpha, \beta \in \mathbb{R}_{>0} , \alpha + \beta = 1 : f(\alpha x+ \beta y) \leq \alpha f(x) + \beta f(y)$.

Proof of absolute value function is convex:

Let x, y $\in \mathbb{R}$, let $\alpha$, $\beta$ $\in \mathbb{R}_{>0}$ where $\alpha + \beta = 1$.

$f(\alpha x + \beta y) = |\alpha x + \beta y| \leq |\alpha x| + |\beta y | = |\alpha| |x| + |\beta| |y| = \alpha |x| + \beta |y| = \alpha f(x) + \beta f(y)$

therefore, $f(\alpha x + \beta y) \leq \alpha f(x) + \beta f(y)$, proving the convexity.



##Q4-7.28
###Prove that the elastic net objective function is convex using the results from the previous two exercises.

Given a data matrix X and response vector y, the elastic net is defined as the solution to the optimization task
$$arg min \left \{ \frac1{2n} ||y-Xb||^2_2 + \lambda((1-\alpha) \frac1{2} ||b||^2_2 + \alpha ||b||_1) \right\}$$
for any values $\lambda >0$ and $\alpha \in [0,1]$.

Based on the result from part(b) that the absolute value function is convex, we can say the terms $\frac1{2n} ||y-Xb||^2_2$,  $\lambda((1-\alpha) \frac1{2} ||b||^2_2$ and $\alpha ||b||_1$ are all convex; based on the result from part(a), we can say the sum of the above 3 terms are also convex. Therefore, the elastic net( as the argmin of the sum of 3 absolute values) is also convex.


##Q5-7.31
###Find the KKT conditions for glmnet when $\alpha = 1$  and implement a `lasso_reg_with_screening` function that takes an $\alpha$ parameter. 

Use glmnet to get all the coefficients and sum to 0. Use KKT to check any violation.
```{r}
lasso_reg_with_screening <- function(X, y, b, lambda)
{
resids <- y - X %*% b
s <- apply(X, 2, function(xj) crossprod(xj, resids)) / lambda / nrow(X)
# Return a vector indicating where the KKT conditions have been violated by the variables that are currently zero.
(b == 0) & (abs(s) >= 1)
}
```


